---
output: 
  github_document:
    toc: FALSE
    toc_depth: 3
---

# 👋 Thanks for dropping by!


Stuff I do:

- [x]  Descriptive Statistics, Statistical Modeling, Machine Learning (with emphasis on viz)
- [x]  Visualization (ggplot2, plotnine, Tableau), Table Generation (tidypivot, gt, greattables, flextable)
- [x]  Dashboarding (Experience with Shiny, powerBI, Tableau)
- [x]  Research Reporting (Scholarly, Institutional; Rmarkdown, Quarto)
- [x]  Workflow/Pipeline Management (tidyverse, base R, SQL)
- [x]  Research Software Engineering (experienced R package maintainer, > 5 years)
- [x]  Statistical Consulting
<!-- - [x]  Technical Communication (oral and written) -->
- [x]  Data Science Education (excellence in class, and world wide)
- [x]  Version Control, Collaborative Workflows, Git, Github, Github Actions
- [x]  Productive, Professional, and Friendly Working Relationships with Collaborators with diverse technical and subject area backgrounds
- [x]  Tidyverse Fluency (dplyr, ggplot2, tidyr, purrr, etc)
- [x]  R Packaging Expertise (workflow tools, documentation, testing)
- [x]  Visualization Frameworks (spatial, network, grammar of graphics)
- [x]  Contributor to popular open source libraries (ggplot2, plotnine)
- [x]  Author, maintainer of popular libraries (flipbookr, ggcirclepack)
- [x]  Experience with CRAN submission process and maintanance (flipbookr, tidytitanic)

<!-- Bachelor's degree in software development or a directly related field from an accredited institution. -->

<!-- Three (3) years of software development, software or system engineering, bioinformatics, or directly related IT professional experience. -->

<!-- Two (2) years of experience in data analysis using R with common analysis utilities (Tidyverse, dplyr, etc.). -->

<!-- Two (2) years of experience with manipulating large datasets to transform, profile, sanitize, explore, analyze, and present information in a reproducible way. -->

<!-- Two (2) years of experience with version control, preferably with Git -->

<!-- Experience developing and deploying reusable and robust R packages, including familiarity with testing frameworks (testthat) and documentation generation (roxygen2). -->


<!-- A combination of education and related technical/military/paraprofessional experience may be substituted for a bachelor’s degree on a year for year basis. An advanced degree (Masters or Doctorate) may be substituted for experience on a year for year basis if the degree is in a field of study directly related to the work assignment. -->


<!-- - [x]  Reproducible computing environments, containerization -->

```{r, include = F}
library(tidyverse)
knitr::opts_chunk$set(warning = F, message = F, echo = F)

```



I work on tools that make data analytics more fluid and intuitive and on tools which expose this fluidity and elegance. These tools also allow us to better record 'conversations' with data and workflows.



<!-- ```{r} -->


<!-- ``` -->


<!-- exported functions: `r getNamespaceExports("ggplyr")` -->

<!-- tidypivot - -->

<!-- exported functions: `getNamespaceExports("tidypivot")` -->

<!-- ggdims - -->

<!-- exported functions: `r getNamespaceExports("ggdims")` -->

<!-- ggregions - `r getNamespaceExports("ggregions")` -->





<!-- statexpress - `r getNamespaceExports("statexpress")` -->


<!-- Record conversations: -->

<!-- ggprop.test - `r getNamespaceExports("ggprop.test")` -->



<!-- Expose this fluidity and elegance: -->

<!-- flipbookr - `r getNamespaceExports("flipbookr")` -->

<!-- ggram - -->

<!-- Education: -->

<!-- easy-geom-recipes -->

<!-- easy-geom-recipes-python -->

<!-- easy.geom.recipes.package -->


<!-- knitrExpress -->



My popular projects include [flipbookr](https://github.com/EvaMaeRey/flipbookr) (on CRAN and maintained since 2020, with \>20000 downloads and used across disciplines) and ggplot2 extensions like [ggcalendar](https://evamaerey.github.io/ggcalendar/), [ggcirclepack](https://github.com/EvaMaeRey/ggcirclepack) and others.

Professionally, I've worked in higher education and government where my work involved analytics, teaching, and policy implementation.

I'm interested in lightening cognative load, first, when interpreting data visualizations and, second, when writing and reading code used to build data visualizations. I believe intuitiveness of plot composition tools (coding and gui interfaces) often translates to more effective visualization. With logical, easy-to-use tools, we're better positioned to build compelling, easy-to-interpret visualizations, rather than stopping at 'good enough'. For these reasons, I'm a big fan of the elegant and intuitive grammar of graphics visualization frameworks.

My technical expertise is in ggplot2 and my current focus is on extension and supporting extenders. I co-founded and organize the [ggplot2 extenders club](https://ggplot2-extenders.github.io/ggplot-extension-club/)...

, have created ['easy geom recipes'](https://evamaerey.github.io/easy-geom-recipes/), am writing [ggplot2 extension cookbook](https://github.com/EvaMaeRey/ggplot2-extension-cookbook), and developing ['express' methodologies for extension](https://github.com/EvaMaeRey/ggexpress) (why shouldn't everyone be using extension and why shouldn't we be using them even on an *ad hoc* basis?).

I've studied some of these new educational materials via [survey response and focus groups](https://evamaerey.github.io/easy-geom-recipes/survey_results_summary.html). For more on the motivation for these efforts, see ['everyday ggplot2 extension'](https://evamaerey.github.io/everyday_ggplot2_extension/).

I'm interested in the transformational effects that access to tailored, principled data visualization tools can have on analytic and teaching spaces. I work on greater accessibility for analysts, research, and students to craft tool suited to their particular data challenges.

Previously, I had a greater focus on illuminating the grammar of *base* ggplot2, creating materials like [a ggplot2 grammar guide](https://evamaerey.github.io/ggplot2_grammar_guide/about) and [the ggplot2 flipbook](https://evamaerey.github.io/ggplot_flipbook/ggplot_flipbook_xaringan.html#1).

I am also especially interested in entry points to R package writing, working on resources like a ['companion guide](https://evamaerey.github.io/package_in_20_minutes/package_in_20_minutes) to to Jim Hester's talk how to write and R package in 20 minutes' (2000) I have also looked comparatively at literate package writing tools of [{fusen}](https://thinkr-open.github.io/fusen/), [{litr}](https://jacobbien.github.io/litr-project/), and my own 'readme-to-package' approach, including via an coordinating a virtual meeting with the authors at RLadies Denver, [March 2024 meeting](meetup.com/rladies-denver/events/299879858/?eventOrigin=group_past_events). The [note package (2025)](https://github.com/musician-tools/note) uses my [{knitrExtra}](https://github.com/EvaMaeRey/knitrExtra) package to create the note package (a lightly rewritten version of what Jim Hester presented) from within a README.

Some projects that I'm especially excited about at the moment are:



```{r, fig.width=10, fig.height=6, dpi = 300, eval = T}
new_style = ggram:::specify_code_plot_style(vline_color = "transparent", accent = NULL, hline_color = "grey90",
                                            paper_color = "cornsilk1" |> alpha(.4), highlight_colors = c("grey90" |> alpha(.4), "orange" |> alpha(.3) ))

'library(ggdims)

iris |> 
  ggplot() + 
  aes(dims = #<<
        dims( #<<
        Sepal.Length:Petal.Width #<<
        ), #<<
      fill = Species) + 
  geom_tsne()  #<<' |>
ggram::ggram(code = _ , "ggdims API EvaMaeRey/ggdims", widths = c(2, 1.5), code_style_args = new_style,
             caption = "github.com/EvaMaeRey/ggdims")

```



```{r, eval = F}
library(ggplot2)
# panel_gray_gradient <- 
  

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scales::pal_grey()(10) |> 
  grid::linearGradient() |> 
  theme(panel.background = _)

gray_gradient <- grid::linearGradient(scales::pal_grey()(10))

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  theme(panel.background = element_rect(fill = gray_gradient))


library(ggplot2)

my_gradient <- c("lightblue", #<<
                 "pink") |> #<<
  grid::linearGradient() #<<

ggplot(cars) +
  aes(speed, dist) + 
  geom_point(shape = "🎈",
             size =7) +
  theme(
    panel.background = 
      element_rect(
        fill = my_gradient) #<<
        )

ggram::ggram("The sky's the limit!", subtitle = "R 4.1.0 started supporting patterns and gradients which meant new possibilities for using ggplot2\nincluding fill gradients")


ggplot(cars) +
  aes(speed, dist) + 
  geom_point(shape = "🎈",
             size =7) +
  theme_panel.background()


```


