---
title: National Anthems’ Sentiment Scores, Mapped and Interactive
author: Evangeline Reynolds
date: '2018-08-04'
slug: national-anthems-sentiment-scores-mapped-and-interactive
categories:
  - text analysis
  - maps
  - nationalism
tags:
  - text analysis
  - maps
  - nationalism
header:
  caption: ''
  image: ''
---

# Background

This post, as indicated in the title, is about an interactive mapping of sentiment scores calculated for national anthems.  Text analysis is of growing interest for political researchers, and I count myself among the interested!  The interactive plot at the end of the post is, I think, an ideal introduction sentiment analysis.  It highlights opportunities, and also, perhaps, some pitfalls.     

I want to give quick background on how this happens to be in my blog, to give plenty of credit where it is due.  I came across the following tweet of some folks that I met at the Lorentz Workshop 'Empirical Research on International Organizations' held in the Netherlands in January, 2018.    

```{r echo=FALSE}
blogdown::shortcode('tweet', '996874882515337216')
```

What a great idea!  The national anthems are a small, but tremendously interesting corpus.  Most people are intimately familiar with at least one document (their own country's anthem), and will be interested in the resulting sentiment score and how it compares to other countries'.  And the visualizing an outcome like a sentiment score on a map is an appealing way to communicate results.  

By the way, Marinko is a lecturer and Jarek is a Assistant Professor at the University of Leiden.  Jarek was also instrumental in bringing us together at the conference, as one of the proposers and organizers. Thanks for the many emails on logistics, Jarek! (Thinking of proposing a Lorentz workshop yourself? You don't have to be local. [Here](https://www.lorentzcenter.nl/proposals.php) is the link to find out details on organizing a Lorentz workshop.)

Back to sentiment analysis. To briefly describe the work up to this point, Marinko and Jarek gathered English translations of the anthem texts from a few of websites, wikipedia, lyricsondemand.com, and nationalanthems.info. 

Then they performed a sentiment scoring using the software Sentistrength; [here](http://sentistrength.wlv.ac.uk/) you can read more about the algorithm, performance, and how it has been used.  Sentistrength returns two sentiment strengths: both strength of positive and of negative sentiments.  Here are distibutions on both of those measures for the anthems in their dataset.  

```{r, echo = F, fig.show='hold', fig.width=2.5, fig.height=2, message = F, warning=F}
load("../../../side_projects/anthems_sentiment_interactive/environment_anthems.Rdata")

library(ggplot2)
library(tidyverse)

ggplot(df) +
  aes(x = Positivesentiment) +
  geom_bar(alpha = .2, fill = "blue") +
  scale_y_continuous(limits = c(0, 150)) +
  theme_bw() + labs(title = "Positive Sentiment", x = "score")

ggplot(df) +
  aes(x = abs(Negativesentiment)) +
  geom_bar(alpha = .2, fill = "blue") +
  scale_y_continuous(limits = c(0, 150)) +
  theme_bw() + labs(title = "Negative Sentiment", x = "score (magnitude)")

```
You see that there is less variation on the positive than the negative sentiment scores.  






Marinko and Jarek mapped the *difference in absolute value* for these two sentiment scores, which is also what we do for the interactive version.  Here is that distribution: 

```{r, echo=F, warning=F}
ggplot(df) +
  aes(x = Positivesentiment - abs(Negativesentiment)) +
  geom_bar(alpha = .2, fill = "blue") +
  scale_y_continuous(limits = c(0, 150)) +
  theme_bw() + labs(title = "Difference in Magnitudes\nPositive minus Negative", x = "score (magnitude)")

```
The extreme cases are Samoa and Tuvalu where the difference is 3 (much more positive than negative sentiment), and Cyprus, Guatemala and Peru where the difference is -3.


# Interactive version

So after seeing their initial work, I proposed an interactive version so that people that are new to text-as-data analyses could see the exact texts that produced the variation in scores.  Marinko and Jarek agreed.  We used r + plotly to produce this final (but still draft) interactive version; when you mouse over a country, the value for the magnitude difference (positive - negative) is displayed, as well as the anthem text!  For best results view on a full size tablet or computer.

Enjoy exploring!

# Plot is not rendering, try again later. 

```{r, echo = F, fig.width=15, fig.height=8, eval = F}
p
```

  

# Lessons?

We think that at least one important lesson can be demonstrated with this tool:  interpretation should be cautious and informed by a reading of a good sample of texts at different sentiment levels.  When survaying the map, we see some regions where negative scores are greater than the positive scores, for example the Americas.  We might be tempted to conclude that the outlook in the Americas is pessimistic based on their anthems.  However, much of the negative language is due to discussion of overcoming oppression, reflecting colonial histories; there may be a large share of negative words in what are ultimately very hopeful and inspiring texts.  






