---
title: What the flipbook?!
author: ~
date: '2020-03-08'
slug: introducing-flipbookr
categories: []
tags: [flipbooks, flipbookr, communication, slowggplot2]
subtitle: ''
summary: ''
authors: []
lastmod: '2020-03-08T19:12:13-06:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

At the end of January, I was lucky enough to present at the RStudio 2020 Conference about flipbooks and {flipbookr}.  

About a year earlier, I'd tweeted about [*the ggplot flipbook*](https://evamaerey.github.io/ggplot_flipbook/ggplot_flipbook_xaringan.html) a collection of data visualizations I'd made  ---  presenting the *builds* of the code and visualizations.  For my standards, the tweet got an *amazing* response and was widely shared.  

Heading to the conference, I felt like the word was out about flipbooks and I'd mostly be presenting about {flipbookr}, the new package I was developing to help people make their own flipbooks. But actually, many people that I met had never heard of or seen any flipbooks.  

So, maybe you *too* haven't heard of flipbooks.  What are they?  Well, flipbooks present code step-by-step and side-by-side with output.  It's probably easiest to understand by looking at an example.  A flipbook is embedded below.  Click in the box and then use the arrow keys to navigate through it: 

```{r, echo = F}
knitr::include_url("https://evamaerey.github.io/flipbooks/racing_bars/racing_barcharts.html")
```

The idea of flipbooks is to deliver "readers" insights about behavior of specific functions in the context of a larger project. The "reader" can isolate unfamiliar functions and toggle between output states before and after a function is used --- seeing how that function affects the outcome.

# Where to get informed

Now the whole flipbook project has a new [home page](https://evamaerey.github.io/flipbooks/about){target="_blank"} where I've tried to provide information on all things flipbooks.  It includes links to many examples, info on how to build your own flipbooks with {flipbookr}, details about how {flipbookr} works, and a bit on the flipbooks origin story.  Please check it out if you want to know more!  


```{r, echo = F}
knitr::include_url("https://evamaerey.github.io/flipbooks/about")
```



## The flipbook building package, {flipbookr}

The package {flipbookr} helps creators build flipbooks by allowing them to create *a single pipeline of code as an input*, and then apply a single function which:

1. parses their source pipeline
2. creates partial pipeline builds
3. delivers the partial pipelines and output to slides in slide show tool (xaringan)

Bam! You'll have a tool to precisely communicate about code pipelines with colleagues, students, or "your future self".  More details about how exactly to get started are also at the [flipbooks home page](https://evamaerey.github.io/flipbooks/about).


<!-- The intent is that {flipbookr} will help creators painlessly build flipbooks for reference or to communicate with colleagues or students.  In the long run, I'm personally hopeful that {flipbookr} will also see much student use.  Students might build their own bespoke flipbooks --- portfolios of data visualizations and data manipulation for personal reference.     -->


<!-- # Getting started with {flipbookr} -->

<!-- Here are the steps to getting started with {flipbookr} in RStudio. -->

<!-- 1. `install.packages("xaringan")`  *{xaringan} is a tool that allows you to create remark.js slide shows from an Rmarkdown file -- so you are able to embed code and output therein* -->

<!-- 2.  -->
<!--     a. `install.packages("remotes")` *since {flipbookr} is just a development package, we need to install it from github* -->
<!--     b. `remotes::install_github("flipbookr")`  -->
<!-- 3. **Restart R and RStudio** -->
<!-- 4. **Files -> New File -> Rmarkdonw -> from template -> A Minimal Flipbook ** *This template will be made available to you upon installation of the flipbookr package* -->
<!-- 5. **"knit" the template** *the template contains a lot of examples about the different modes for using {flipbookr}* -->
<!-- 6. modify the template to make it your own! -->


# Feedback for {flipbookr}?

At the RStudio conference I made the claim:

> "With {flipbookr}, you can build a flipbook in 5 minutes".  

Such promises are enticing but in my experience rarely quite true! The *you* in these types of claims never seem to be referring to *me*, at least.  It is a really impersonal, generic *you*, that actually refers to the person delivering demo.  What would ring a little more true is "Let me show you how *I* can build a flipbook in 5 minutes."    

<!-- > "*If* you have the package successfully installed, *and if* you are really concentrated  *and if* follow the demo very carefully  *and if* you do not deviate from the instructions *then* building your first flipbook *might* take you 5 minutes."   -->

Yup, saying "5 minutes" was probably a little disingenuous.  I don't feel too bad though --- these promises have worked on me --- getting me to try a package building and blogdown tutorials for example.  Just took me, you know, 4 to 7 times longer than promised that first time through!  But I learned the new tools and skills!

And so now **I'd love to know what the *real life* {flipbookr} experience is**!    Have you tried to build a flipbook?  How much of an exaggeration was "5 minutes"?    How was installation?   Did you find it easy/difficult/confusing?  Were you successful? Did you maybe give up?  Was there some particular code that you tried to "flipbook" that threw an error instead? Would you be willing to share your flipbook itself? 

The easiest way to reach me is probably on Twitter @EvaMaeRey or submitting an issue on Github --- please say it nicely if you've got gripes! Or, if you'd like to make your comments privately, just "at" me and we can arrange another way!  Thanks.







