---
title: Embedding a flipbook mini in an html document
author: ~
date: '2020-03-21'
slug: embedding-flipbook-mini-in-html-document
categories: []
tags: []
subtitle: ''
summary: ''
authors: []
lastmod: '2020-03-21T18:52:53-06:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---


```{r setup, include=F}
knitr::opts_chunk$set(warning = F, message = F)
library(tidyverse)
```


Xaringan is a fantastic rmarkdown presentation tool from Yihui Xie which uses remark.js as its platform. (Yihui Xie has a lot to say about the greatness of remark.js itself in a [blogpost](https://yihui.org/en/2017/08/why-xaringan-remark-js/){target="_blank"} --- worth a read)  
One shortcoming of the format is that the slide shows are not self-contained. So a presentation is usually a whole collection of files.   This is obviously true for [*flipbooks*]() created with the {xaringan} platform. 

Having so many files might feel a bit overwhelming, especially when it comes time to share your work. So some time ago, I had the idea of working on something self-contained: a flipbook mini (yup, like "ipod mini"). What's been worked out is using all the parsing code from {flipbookr} and then plotting the code itself in a ggplot, and producing the plot output.  Then using {cowplot} the images are pasted side-by-side together.  This is done for each partial build of the code.  The resultant side-by-side figures are then combined into a gif for the movie effect.  

<blockquote class="twitter-tweet"><p lang="en" dir="ltr">And checking out new magic in <a href="https://twitter.com/hashtag/ggforce?src=hash&amp;ref_src=twsrc%5Etfw">#ggforce</a>. Adapted from <a href="https://twitter.com/thomasp85?ref_src=twsrc%5Etfw">@thomasp85</a>&#39;s examples: <a href="https://t.co/aYMKuV7uIn">https://t.co/aYMKuV7uIn</a> <a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a> <a href="https://t.co/LNpci0h8OL">pic.twitter.com/LNpci0h8OL</a></p>&mdash; Gina Reynolds (@EvaMaeRey) <a href="https://twitter.com/EvaMaeRey/status/1162680699834556418?ref_src=twsrc%5Etfw">August 17, 2019</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>


At the RStudio conference someone asked if flipbooks can be embedded in blog posts or other html documents. By conference time, I thought I'd have the mini to demo --- but I didn't, as a lot of work needed to be done on fundamental {flipbookr} internal functions.  I also didn't think of directly embedding into html.  But now the embed too is basically worked out.  After the gif is saved out with flipbookr::chunk_create_gif_flipbook(), I can pull it back in to embed it (this is done with knitr::include_graphics()). The {flipookr} function used to get all this done is `chunk_gif_flipbook_embed()`.  Sadly, this doesn't work in blogdown!  

All of this feels a little awkward so if there are better ways of going about it, do let me know! Check out the details of the implementation [here](https://github.com/EvaMaeRey/flipbookr/blob/master/R/mini.R).

So, first you'll have an input code chunk that you'll refer to by name (I'm calling it cars_plot, and I have echo set to true, but you can do echo = F): 

```{r cars_plot, echo = T, eval = T}
ggplot(data = cars) +
  aes(x = speed) +
  aes(y = dist) +
  geom_point(shape = 21, 
             alpha = .75,
             color = "white") +
  aes(fill = dist) +
  aes(size = speed) +
  scale_fill_viridis_c(end = .9, 
                       option = "magma") +
  theme_minimal() +
  theme(plot.background = 
          element_rect(fill = "grey90"))
```

Then you can embed the flipbook mini as follows (again you might want to set echo to false in the code chunk options):

```{r}
library(flipbookr)
chunk_gif_flipbook_embed(chunk_name = "cars_plot", 
                         display_type = "output")
```


You might of course want to show the code and the output side-by-side -- the classic flipbook. 

```{r}
chunk_gif_flipbook_embed(chunk_name = "cars_plot", 
                         display_type = "both")
```

# Limitations!

The flipbook mini function  works with plot outputs *only* for the time being.  

Also, Claus Wilke has noted that the flipbooks aspect ratio control needs some work -- which is really apparent in the flipbook mini and is also an issue on the {xaringan} platform.  This has been duely noted and is logged as an issue in the development package --- help welcome! 

# More ways of embedding...

Another way you can share flipbooks is using the knitr function include_url() within a code chunk.  I think I learned this from Emi Tanaka initially!

```{r}
knitr::include_url("https://evamaerey.github.io/tidytuesday_walk_through/tidytuesday_highlights.html")
```




